Atomify - a live LAMMPS visualizer

### 2017-06-16 ms.tex
Added info on plotting and dashboard

Any operations that is applied to the system in a LAMMPS timestep or
minimization step is called a ``fix''.
These fixes are defined in the LAMMPS script using the ``fix'' keyword.
Atomify registers all defined fixes and lists these in the dashboard.

TODO: Time evolution of quantity \\
TODO: 2d ave/chunk binning \\

In LAMMPS, computes are physical quantities calculated based on the running
simulation, such as energy, temperature and stress.
% TODO better word for measures?
These are typically macroscopic measures derived from the microscopic
interactions between the atoms.
These can be plotted as a function of time and visualized while the 
simulation is running.
Quantities that are measured per atom can be plotted in a histogram.

Atomify provides live visualization of the compute values using various plotting
techniques.
For instance, the temperature of the simulated system can be plotted as a
function of time as seen in figure TODO.
% TODO add figure

@@

@@

Atomify is built upon Qt.
All GUI is defined in QML whereas the backend of the program communicating with
LAMMPS is written in C++.
High-performance visualization is achieved by implementing billboards in Qt3D.
Simply put, Atomify runs an instance of a LAMMPS object enabling the user to
visualize the current state of LAMMPS and interact with LAMMPS while the
simulation is running.

@@

Simulations of atoms are often visualized using spheres as atoms and cylinders.
The spheres represent the atoms while the cylinders represent the bonds between
the atoms.
This way, the molecular structure is easy to understand and interpret.
In many visualization tools, the spheres and cylinders are built up by many
triangles connected so they form the geometrical object of interest.
One problem with this rendering technique is that each sphere needs many,
often hundreds, of triangles in order to look spherical.
A common technique to overcome this problem is to use billboards.
This significantly reduces the number of vertices sent from the CPU to the GPU,
which in turn improves rendering performance.
A sphere may consist of an arbitrary number of vertices,
depending on the level of detail needed.
A billboard will on the other hand consist of only 4 vertices and can be
represented as a point on the CPU which is then expanded on the GPU using shader
programming.
By doing proper shading, the billboards can mimick the spheres and cylinders so
they look realistic, see Figure \ref{fig:final_billboards}.

@@

\section{Case study (Anders)}
% Rerun on large dataset

>>>>

Any operation that is applied to the system in a LAMMPS timestep or minimization
step is called a ``fix''.
These fixes are defined in the LAMMPS script using the ``fix'' keyword.
Atomify registers all defined fixes and lists these in the dashboard.

In LAMMPS, computes are physical quantities calculated based on the running
simulation, such as energy, temperature and stress.
% TODO better word for measures?
These are typically macroscopic measures derived from the microscopic
interactions between the atoms.
Computes and variables are shown in the Atomify dashboard,
along with information about the regions, groups and fixes as shown in figure
\ref{fig:dashboard}.

@@

Computes can be plotted as a function of time and visualized while the 
simulation is running.
Quantities that are measured per atom can be plotted in a histogram or a 2D
average for each region.
Atomify provides live visualization of the compute values using various plotting
techniques.
For instance, the temperature of the simulated system can be plotted as a
function of time as seen in figure \ref{fig:temperature}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/temperature.png}
	\caption{Atomify can visualize computes defined in LAMMPS. 
    Here the temperature is plotted as a function of time.}
	\label{fig:temperature}
\end{figure}

@@

Atomify runs a LAMMPS instance in the background that the user can issue 
commands to and visualizes the current atom positions and the LAMMPS state.
The user can interact with LAMMPS while the simulation is running.
Atomify is written in C++ and QML using the Qt framework.
The GUI is developed in QML and LAMMPS communication,
computations, and rendering is mostly defined in C++.
High-performance visualization is implemented as billboards in Qt3D.

@@

Simulations of atoms are often visualized using spheres as atoms and cylinders.
The spheres represent the atoms while the cylinders represent the bonds between
the atoms.
This way, the molecular structure is easy to understand and interpret.
In many visualization tools, the spheres and cylinders are built up by many
triangles connected so they form the geometrical object of interest.
One problem with this rendering technique is that each sphere needs many,
often hundreds, of triangles in order to look spherical.
A common technique to overcome this problem is to use billboards.

Billboards are plane objects that are always facing the camera (their normal
vector always points from the billboard center towards the camera center).
Drawing a filled circle on the billboard is a simple approximation to rendering
spheres, but this approach suffers from unrealistic distortions from the field
of view and does not make it possible to render realistic lighting.
The more robust approach, which we have chosen,
is to cast rays from the camera to the plane and calculate the sphere
intersection of each ray.
The normal vector can then be found by subtracting the intersection point from
the sphere center.

The technique of using billboards with only four vertices instead of a mesh
approximating a sphere with many more vertices is often referred to as
\emph{imposter spheres} in the literature.
% TODO add reference to literature explaining the details of rendering spheres with imposter billboards
This technique significantly reduces the number of vertices sent from the CPU to
the GPU, which in turn improves rendering performance.
A sphere may consist of an arbitrary number of vertices,
depending on the level of detail needed.
A billboard will on the other hand consist of only 4 vertices and can be
represented as a point on the CPU which is then expanded on the GPU using shader
programming.
By doing proper shading, the billboards can mimick the spheres and cylinders so
they look realistic, see Figure \ref{fig:final_billboards}.

@@

\section{Case study (Anders)}

% TODO Write about case study
